---
title: "stats101a_hw6"
author: "Takao"
date: "2/14/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### (a)

```{r}
matXtX <- matrix(c(1.17991, -7.30982e-3, 7.3006e-4, -7.30982e-3, 7.9799e-5, -1.2371e-4, 7.3006e-4, -1.23713e-4, 4.6576e-4), nrow = 3)
matXty <- matrix(c(220, 36768, 9965), nrow = 3)
matXtX %*% matXty
```

The regression coefficient in the model specified is `r matXtX %*% matXty`.
y hat = 0.093089(x1) + 0.253232(x2) -1.91221


### (b)

```{r}
yhat1 <- 0.093089*200 + 0.253232*50 -1.91221
yhat1
```

The predicted value of the absorption index y when x1 = 200 and x2 = 50 is `r yhat1`.

### (c)

```{r}
var1 <- 1 * matXtX
var1
```

The variances of the estimated coefficients are `r var1`

## Problem 2

```{r}
electric_data <- read.csv("electric.csv", header = TRUE)
head(electric_data)
```

### (a)

```{r}
mlr.model1 <- lm(y~x1 + x2 + x3 + x4, data = electric_data)
mlr.model1
```
Estimated regression coefficients: 
x1: 0.7573
x2: 7.5188
x3: 2.4831
x4: -0.4811
intercept: -123.1312
y hat = 0.7573(x1) + 7.5188(x2) + 2.4831(x3) - 0.4811(x4) - 123.1312

### (b)

Estimated regression coefficient:
x2: 7.5188
The electric power consumed each month by a chemical plant increases 7.5188 Megawatts per hour (MWh) when the number of days in the month increases by 1 day, when all other factors that influence the electric power consumption are fixed.

x4: -0.4811
The electric power consumed each month by a chemical plant decreases 0.4811 Megawatts per hour (MWh) when the product produced increases by 1 ton, when all other factors that influence the electric power consumption are fixed.


### (c)

```{r}
e <- mlr.model1$residuals # Extract the residuals.
n <- length(e) 
p <- 4 
est.variance1 <- sum(e^2)/(n- p - 1)
print(est.variance1) 
```
Estimation of the error variance is `r est.variance1`. 

### (d)

```{r}
summary(mlr.model1)
```
The standard errors of the regression coefficients:
x1: 0.2791
x2: 4.0101
x3: 1.8094
x4: 0.5552
intercept: 157.2561

As we can see, the standard errors of the regrssion coefficients are fairly distinct, which indicates that not all of the model parameters estimated with the same precision.


### (e)

```{r}
# y hat = 0.7573(x1) + 7.5188(x2) + 2.4831(x3) - 0.4811(x4) - 123.1312

powerconsumption <- 0.7573*75 + 7.5188*24 + 2.4831*90 - 0.4811*98 - 123.1312
powerconsumption
```

The predicted power consumption for a month in which x1 = 75 degrees farenheit, x2 = 24 days, x3 = 90 percent, and x4 = 98 tons is `r powerconsumption` MWh. 

### (f)

```{r}
summary(mlr.model1)
```

At a level of $\alpha = 0.05$, we draw the following conclusions from the t-tests:

1. The average ambient temperature has a significant association with the electric power consumed each month by a chemical plant, when the predictors number of days in the month, product purity, and tons of product produced are in the model.
2. The number of days in the month does not have a significant association with the electric power consumed each month by a chemical plant, when the predictors ambient temperature, product purity, and tons of product produced are in the model.
3. The average product purity does not have a significant association with the electric power consumed each month by a chemical plant, when the predictors ambient temperature, number of days in the month, and tons of product produced are in the model.
4. The tons of product produced does not have a significant association with the electric power consumed each month by a chemical plant, when the predictors, ambient temperature, number of days in the month, and product purity are in the model. 

### (g)

```{r}
confint(mlr.model1, level = 0.95)

```

Confidence Intervals:
x1: (0.09734663, 1.4172315)
x2: (-1.96364640, 17.0012143)
x3: (-1.79543852, 6.7615956)
x4: (-1.79391356, 0.8316431)
Intercept: (-494.98273363, 248.7202411)

As we can see in the various confidence interval, x1 is the only CI that does not contain the value "0" in the interval, which indicates validation to the conclusion we made in part (f). 





