---
title: "stats101A_hw1"
author: "Takao"
date: "1/8/2022"
output:
  html_document:
    df_print: paged
---

## 3

### (a)

NULL Hypothesis: mu = 225 hours (The mean repair time is 225 hours)

Althernative Hypothesis: mu > 225 hours (The mean repair time exceeds 225 hours)

### (b)
```{r}
repair <- c(159, 280, 101, 212, 224, 379, 179, 264, 222, 362, 168, 250, 149, 260, 485, 170)

t.test(repair - 225, alternative = "greater")
```

Since the p-value is greater than the significance level of 0.05, we fail to reject the null hypothesis. There is no significance evidence thatthe mean repair time exceeds 225 hours


### (c)

```{r}
t.test(repair - 225, alternative = "greater")
```

The P-value for the test is 0.257.

###(d)

```{r}
t.test(repair - 225)
lowerv <- 225 - 36.107031
upperv <- 225 + 69.10731
confidenceinterval <- c(lowerv, upperv)
confidenceinterval
```

The 95 percent confidence interval on the mean repair time is `r confidenceinterval`.
We are 95 percent confident that the mean repair time for the specific electronic instrument is between 188.8930 hours and 294.1073 hours. 
