---
title: "stats101A_hw3"
author: "Takao"
date: "1/19/2022"
output:
  pdf_document: default
  html_document: default
---

## 1

```{r}
football_data <- read.csv("football.csv")
head(football_data)
```

### (a)

The response variable is the number of games won or Games.Won
The predictor variable is the number of yards gained rushing by opponents or Opponent.Rushing.Yards


### (b)

```{r}
slr.model1 <- lm(Games.Won ~ Opponent.Rushing.Yards, data = football_data)
print(slr.model1)
b.coeff <- coefficients(slr.model1)

plot(football_data$Games.Won ~ football_data$Opponent.Rushing.Yards, xlab = "# of yards gained rushing by opponents", ylab = "# of games won")

abline(a = b.coeff[1], b = b.coeff[2])

b.coeff
```
Intercept = 21.788251
In the 1976 National Football League, we predict the number of games won (per 14-game season) by a team is 21.788251 games when the number of yards gained rushing by opponents is 0 yards. 

Slope = -0.0070251
In the 1976 National Football League, we predict that the number of games won (per 14-game season) by a team will decrease by 0.0070251 games when the number of yards gained rushing by opponents increases by 1 yard. 


### (c)

```{r}
anova(slr.model1)
```

The ANOVA table shows a p-value of 0.000007381. Based on a significance level of $\alpha = 0.05$, we have significance evidence to reject the null hypothesis. since 0.000007381 < 0.05. There is sufficient evidence that the number of yards gained rushing by opponents has an effect on the number of games won (per 14-game season) by a team. 


### (d)

```{r}
confint(slr.model1, level = 0.95)
```

(-0.009614347, -0.004435854)

If, in repeated random sampling, we construct a large number of confidence intervals, 95\% of those intervals will contain the true parameter value $\beta_1$. 

We are 95\% confident that the interval (-0.009614347, -0.004435854) contains the true value of the slope $\beta_1$.

### (e)

```{r}
summary(slr.model1)
```
The value of $R^2$ is shown at the output "Multiple R-squared:  0.5447" 
This means that the model $Y = \beta_0 + \beta_1 x + \epsilon$ explains 54.47\% of the total variability in the response number of games won (per 14-game season) by a team. 


### (f)

Using the jargon of the problem, we know that this is a mean response so this is a confidence interval, not a prediction interval

```{r}
dataprediction <- data.frame(Opponent.Rushing.Yards = 2000)

predict(slr.model1, dataprediction, interval = "confidence", level = 0.95)
```

(6.765753, 8.710348)

We are 95% confident that the interval (6.765753, 8.710348) contains the true mean number of games won if the opponents' yards rushing is 2000 yards. 



















